Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-08-27 article/article.Rmd
Commit 2024-08-27 10:14

<!-- Aerosol change -->
The clear-sky reference values discussed above are based on the climatological AOD
and WC; hence they cannot describe accurately the long-term variation of
$\text{GHI}_\text{ref}$ due to long-term changes in the two atmospheric constituents,
mainly AOD.  As reported by @Natsis2023, there is a long-term brightening effect in
the GHI data of Thessaloniki for the period 1993 -- 2023, which for clear-sky data
was attributed to long-term changes in aerosol effects.  Therefore, an adjustment of
the $\text{GHI}_\text{ref}$ during the period of study was made using simulations
with the RTM based on the long-term variations of the AOD. As AERONET data start only
in 2003, we used for the period 1993 -- 2005 estimates of changes in AOD at
$340\,\text{nm}$ derived from a collocated Brewer spectrophotometer [@Kazadzis2007]
to calculate the trend in $\text{GHI}_\text{ref}$ due to aerosols during this period.

@@

According to @Kazadzis2007, in the period 1997 -- 2005 the mean AOD at $340\,\text{nm}$
is $0.403$ with a trend of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta$ of $0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).
<!--
Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively, where for 1997 we used the first available WC data
(i.e. 2003) as a proxy, also taken from AERONET, and for a SZA of $55^\circ$ reveal a
trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$. 
-->
Simulations with uvspec for the above Ångström coefficients, assuming constant WC of
$15.6\,\text{mm}$ taken from the Cimel, and for a SZA of $55^\circ$ reveal a trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$.  The SZA of $55^\circ$ was chosen as
representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean clear-sky irradiance,
and finally the trend of $+0.14\,\%$ per year. We applied these two long-term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

>>>>

<!-- Aerosol change -->
The clear-sky reference values discussed above are based on the climatological AOD
and WC; hence they cannot describe accurately the long-term variation of
$\text{GHI}_\text{ref}$ due to long-term changes in the two atmospheric constituents,
mainly AOD.  As reported by @Natsis2023, there is a long-term brightening effect in
the GHI data of Thessaloniki for the period 1993 -- 2023, which for clear-sky data
was attributed to long-term changes in aerosol effects.  Therefore, an adjustment of
the $\text{GHI}_\text{ref}$ during the period of study was made using simulations
with the RTM based on the long-term variations of the AOD. As AERONET data start only
in 2003, we used for the period 1993 -- 2005 estimates of changes in AOD at
$340\,\text{nm}$ derived from a collocated Brewer spectrophotometer [@Kazadzis2007]
to calculate the change in $\text{GHI}_\text{ref}$ due to aerosols during this period.

@@

According to @Kazadzis2007, in the period 1997 -- 2005 the mean AOD at $340\,\text{nm}$
is $0.403$ with a change of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta$ of $0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).
<!--
Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively, where for 1997 we used the first available WC data
(i.e. 2003) as a proxy, also taken from AERONET, and for a SZA of $55^\circ$ reveal a
change of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$. 
-->
Simulations with uvspec for the above Ångström coefficients, assuming constant WC of
$15.6\,\text{mm}$ taken from the Cimel, and for a SZA of $55^\circ$ reveal a change of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$.  The SZA of $55^\circ$ was chosen as
representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean clear-sky irradiance,
and finally the change of $+0.14\,\%$ per year. We applied these two long-term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

### 2024-08-27 article/article.tex
Commit 2024-08-27 06:14
Commit 2024-08-27 10:14

The clear-sky reference values discussed above are based on the
climatological AOD and WC; hence they cannot describe accurately the
long-term variation of \(\text{GHI}_\text{ref}\) due to long-term
changes in the two atmospheric constituents, mainly AOD. As reported by
\citet{Natsis2023}, there is a long-term brightening effect in the GHI
data of Thessaloniki for the period 1993 -- 2023, which for clear-sky
data was attributed to long-term changes in aerosol effects. Therefore,
an adjustment of the \(\text{GHI}_\text{ref}\) during the period of
study was made using simulations with the RTM based on the long-term
variations of the AOD. As AERONET data start only in 2003, we used for
the period 1993 -- 2005 estimates of changes in AOD at
\(340\,\text{nm}\) derived from a collocated Brewer spectrophotometer
\citep{Kazadzis2007} to calculate the trend in \(\text{GHI}_\text{ref}\)
due to aerosols during this period.

According to \citet{Kazadzis2007}, in the period 1997 -- 2005 the mean
AOD at \(340\,\text{nm}\) is \(0.403\) with a trend of
\(-3.8\pm0.93\,\%\) per year, corresponding to a change of \(0.0153\)
per year. Using an Ångström coefficient \(\alpha = 1.6\), this
translates to a change in the Ångström coefficient \(\beta\) of
\(0.00272\) per year (or \(\beta=0.084\) in 1997 and \(\beta=0.059\) in
2005). Simulations with uvspec for the above Ångström coefficients,
assuming constant WC of \(15.6\,\text{mm}\) taken from the Cimel, and
for a SZA of \(55^\circ\) reveal a trend of \(+0.21\,\%\) per year in
\(\text{GHI}_\text{ref}\). The SZA of \(55^\circ\) was chosen as
representative of all days in the year in order to get a rough estimate
of the annually averaged change in clear sky irradiance. For the period
2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET
in a similar simulation scheme to calculate the monthly mean clear-sky
irradiance, and finally the trend of \(+0.14\,\%\) per year. We applied
these two long-term changes (see
Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-example-years-13}

>>>>

The clear-sky reference values discussed above are based on the
climatological AOD and WC; hence they cannot describe accurately the
long-term variation of \(\text{GHI}_\text{ref}\) due to long-term
changes in the two atmospheric constituents, mainly AOD. As reported by
\citet{Natsis2023}, there is a long-term brightening effect in the GHI
data of Thessaloniki for the period 1993 -- 2023, which for clear-sky
data was attributed to long-term changes in aerosol effects. Therefore,
an adjustment of the \(\text{GHI}_\text{ref}\) during the period of
study was made using simulations with the RTM based on the long-term
variations of the AOD. As AERONET data start only in 2003, we used for
the period 1993 -- 2005 estimates of changes in AOD at
\(340\,\text{nm}\) derived from a collocated Brewer spectrophotometer
\citep{Kazadzis2007} to calculate the change in
\(\text{GHI}_\text{ref}\) due to aerosols during this period.

According to \citet{Kazadzis2007}, in the period 1997 -- 2005 the mean
AOD at \(340\,\text{nm}\) is \(0.403\) with a change of
\(-3.8\pm0.93\,\%\) per year, corresponding to a change of \(0.0153\)
per year. Using an Ångström coefficient \(\alpha = 1.6\), this
translates to a change in the Ångström coefficient \(\beta\) of
\(0.00272\) per year (or \(\beta=0.084\) in 1997 and \(\beta=0.059\) in
2005). Simulations with uvspec for the above Ångström coefficients,
assuming constant WC of \(15.6\,\text{mm}\) taken from the Cimel, and
for a SZA of \(55^\circ\) reveal a change of \(+0.21\,\%\) per year in
\(\text{GHI}_\text{ref}\). The SZA of \(55^\circ\) was chosen as
representative of all days in the year in order to get a rough estimate
of the annually averaged change in clear sky irradiance. For the period
2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET
in a similar simulation scheme to calculate the monthly mean clear-sky
irradiance, and finally the change of \(+0.14\,\%\) per year. We applied
these two long-term changes (see
Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-example-years-12}
